Spatial Analysis

### 2022-01-12 Scripts/spatial_analisis.Rmd
Created partial figures and an alternative Fig 1b

@@

@@

ports_path <- my_path("D","NOAA", name = "PalaciosJ - BSB, SFlounder, Scup by Port 1970-2020 - 2021.10.25.xlsx")

@@

# Check for missing ports
noaa_ports_complete %>% 
  filter(tresh == "75th",
    is.na(lat)) %>% 
  group_by(port_name,state_postal) %>% 
  summarise(n()) %>% 
  View()

# Save data for future analysis
write_csv(noaa_ports_complete,
          my_path("D","Partial", name = "top_ports.csv")
          )

@@

noaa_ports_complete %>% 
  filter(tresh == "75th") %>% 
  ggplot() +
  geom_point(
    aes(
      x = lon,
      y = lat,
      shape = species_name, 
      color = species_name
    ),
    size = 5,
    alpha = 0.5
  )

@@

@@

#Load top port data 
ports_df <- my_path("D","Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "75th")

@@

@@

### NUmbers for papaer

@@

# Load grid of state waters
grid_eez_sw_sf <-  st_read(my_path("D","Partial/grid_sw", name = "grid_eez_sw_sf.shp")) %>%
  select(index,state,geometry) %>% 
  mutate(method = "state_waters",
         state = str_to_sentence(state)) %>% 
  left_join(state_lat)

@@

# Load grid of fishing ports
grid_eez_fp_sf <-  st_read(my_path("D","Partial/grid_fp", name = "grid_eez_fp.shp")) %>%
  select(index,abrev,geometry) %>% 
  mutate(method = "fishing_port"#,
         # state = str_to_sentence(landing_port)
         ) %>% 
  left_join(state_lat)

@@

p_sw <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color =group , fill = group), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45))+
    labs(x = "", y = "", title = "A) State waters approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    geom_sf(aes(color = group),size = 0.1, alpha = 0.3) +
    facet_wrap(~ group) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       # labels = unique(grid_eez_sw_sf$group)) +
                       labels = grid_eez_sw_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

@@

grid_eez_fp_sf <- grid_eez_sw_sf %>% 
  filter(!group %in% grid_eez_fp_sf$group) %>% 
  mutate(index = NA) %>% 
  bind_rows(grid_eez_fp_sf) %>% 
  distinct(.keep_all = T)

p_fp <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(data = subset(grid_eez_fp_sf, index != "NA")) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color = group, fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    # scale_color_manual(values = state_pallet) +
    # scale_fill_manual(values = state_pallet) +
    scale_color_manual(values = c(state_pallet[3:7],state_pallet[9:11])) +
    scale_fill_manual(values = c(state_pallet[3:7],state_pallet[9:11])) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45)) +
    labs(x = "", y = "", title = "B) Fishing ports approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot() +
    geom_sf(data = subset(grid_eez_fp_sf, index = "NA"), color = "white") +
    geom_sf(data = subset(grid_eez_fp_sf, index != "NA"), aes(color = group),size = 0.1, alpha = 0.3) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    facet_wrap(~ group) +
    # geom_point(data = rename(ports_df, group = state_postal), aes(x = lon, y = lat), color = "#E6A0C4") +
    theme(legend.position = "top") +
    # scale_color_manual(values = state_pallet,
    scale_color_manual(values = c(state_pallet[3:7],state_pallet[9:11]),
                       # labels = unique(grid_eez_fp_sf$group)) +
                       labels = grid_eez_fp_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

>>>>

# Save image for manuscript

eez_plot <- ggplot(eez_sf) +
  geom_sf()

name <- my_path("G", "AcrossBoundaries/Results/Figures", name = "eez_map.png")
ggsave(name,
       eez_plot)

@@

# Visual exploration (o.k! )
state_mpa <- st_read("~/OneDrive - UW-Madison/FederalAndStateWaters/FederalAndStateWaters.gdb") %>%
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  mutate(state = str_to_sentence(jurisdicti)) %>% 
  st_transform(crs = 4326) %>%   # for matching projections
  st_simplify(preserveTopology = TRUE, dTolerance = 0.01) %>% 
  left_join(state_lat)

state_mpa$group <- factor(state_mpa$abrev,      # Reordering group factor levels
                               levels = state_lat$abrev)

state_mpa <- ggplot() +
  geom_sf(data = eez_sf,aes(), fill = "grey90") +
  geom_sf(data = land_sf,aes(), fill = "grey50") +
  geom_sf(data = state_mpa, aes(fill = group, color = group)) +
  scale_fill_manual(values = state_pallet) +
  scale_color_manual(values = state_pallet) +
  my_ggtheme_p(leg_pos = "right")

name <- my_path("G", "AcrossBoundaries/Results/Figures", name = "state_sw_map.png")
ggsave(name,
       state_mpa)

@@

ports_path <- my_path("G","AcrossBoundaries/Data/NOAA", name = "PalaciosJ - BSB, SFlounder, Scup by Port 1970-2020 - 2021.10.25.xlsx")

@@

# Check for missing ports (Should be empty!)
noaa_ports_complete %>% 
  filter(tresh == "75th",
    is.na(lat)) %>% 
  group_by(port_name,state_postal) %>% 
  summarise(n()) %>% 
  View()

# Save data for future analysis
# write_csv(noaa_ports_complete,
#           my_path("D","Partial", name = "top_ports.csv")
#           )

@@

port_map <- noaa_ports_complete %>%  
  filter(tresh == "75th")

port_map <- ggplot() +
  geom_sf(data = eez_sf,aes(), fill = "grey90") +
  geom_sf(data = land_sf,aes(), fill = "grey50") +
  geom_point( data = port_map,
    aes(
      x = lon,
      y = lat,
      shape = species_name, 
      color = species_name
    ),
    size = 3,
    alpha = 0.5
  ) +
  my_ggtheme_p(leg_pos = "right")

name <- my_path("G", "AcrossBoundaries/Results/Figures", name = "state_fp_map.png")

ggsave(name,
       port_map)

@@

# ------------------------------ #
# Saving for plot
# ------------------------------ #

state_bf_plot <- state_bf %>% 
  left_join(state_lat)

state_bf_plot$group <- factor(state_bf_plot$abrev, levels = state_lat$abrev)

state_sf_plot <- state_sf %>% 
  left_join(state_lat)

state_sf_plot$group <- factor(state_sf_plot$abrev, levels = state_lat$abrev)

state_bf_map <- ggplot(state_bf_plot) +
  geom_sf(data = eez_sf,aes(), fill = "grey90") +
  geom_sf(data = land_sf,aes(), fill = "grey50") +
  geom_sf(data = state_bf_plot, aes(color = group), fill = NA) +
  geom_sf(data = state_sf_plot, aes(color = group), fill = NA) +
  scale_color_manual(values = state_pallet) +
  my_ggtheme_p(leg_pos = "right")

name <- my_path("G", "AcrossBoundaries/Results/Figures", name = "state_bf_plot_sw.png")

ggsave(name,
       state_bf_map)

@@

#Load top port data 
ports_df <- my_path("G","AcrossBoundaries/Data/Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "75th")

@@

 #------------------------------ #
# Saving for plot
# ------------------------------ #

state_bf_plot_fp <- port_bf %>% 
  rename(abrev = state_postal) %>% 
  left_join(state_lat)

state_bf_plot_fp$group <- factor(state_bf_plot_fp$abrev, levels = state_lat$abrev)

state_bf_map_fp <- ggplot() +
  geom_sf(data = eez_sf,aes(), fill = "grey90") +
  geom_sf(data = land_sf,aes(), fill = "grey50") +
  geom_sf(data = state_bf_plot_fp, aes(color = group), fill = NA) +
  geom_sf(data = state_sf_plot, aes(color = group), fill = NA) +
  scale_color_manual(values = state_pallet) +
  my_ggtheme_p(leg_pos = "right")

name <- my_path("G", "AcrossBoundaries/Results/Figures", name = "state_bf_plot_fp.png")

ggsave(name,
       state_bf_map_fp)

@@

### Numbers for papaer

@@

# Load grid of state waters
grid_eez_sw_sf <-  st_read(my_path("G","AcrossBoundaries/Data/Partial/grid_sw", name = "grid_eez_sw_sf.shp")) %>%
  select(index,state,geometry) %>% 
  mutate(method = "state_waters",
         state = str_to_sentence(state)) %>% 
  left_join(state_lat)

@@

# Load grid of fishing ports
grid_eez_fp_sf <-  st_read(my_path("G","AcrossBoundaries/Data/Partial/grid_fp", name = "grid_eez_fp.shp")) %>%
  select(index,abrev,geometry) %>% 
  mutate(method = "fishing_port"#,
         # state = str_to_sentence(landing_port)
         ) %>% 
  left_join(state_lat)

@@

p_sw <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = eez_sf, aes(), fill = "grey90") + 
    geom_sf(aes(color =group , fill = group), alpha = 0.3) +
    geom_sf(data = land_sf, aes(), fill = "grey50") +
    # ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
    #                                size = 4,
    #                                box.padding = 0.10,
    #                                hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45))+
    labs(x = "", y = "", title = ""),
    # theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = land_sf, aes(), fill = "grey50") +
    geom_sf(aes(color = group),size = 0.1, alpha = 0.3) +
    facet_wrap(~ group) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       # labels = unique(grid_eez_sw_sf$group)) +
                       labels = grid_eez_sw_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

name <- my_path("G", "AcrossBoundaries/Results/Figures", name = "sw_reg_units.png")

ggsave(name,
       p_sw,
       width = 9,
       height = 9)

@@

# Include state with no Regulaotry units in plot
grid_eez_fp_sf <- grid_eez_sw_sf %>% 
  filter(!group %in% grid_eez_fp_sf$group) %>% 
  mutate(index = NA) %>% 
  bind_rows(grid_eez_fp_sf) %>% 
  distinct(.keep_all = T)

p_fp <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(data = subset(grid_eez_fp_sf, index != "NA")) +
    geom_sf(data = eez_sf, aes(), fill = "grey90") + 
    geom_sf(aes(color = group, fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes(), fill = "grey50") +
    # ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
    #                                size = 4,
    #                                box.padding = 0.10,
    #                                hjust = 1) +
    # scale_color_manual(values = state_pallet) +
    # scale_fill_manual(values = state_pallet) +
    scale_color_manual(values = c(state_pallet[3:7],state_pallet[9:11])) +
    scale_fill_manual(values = c(state_pallet[3:7],state_pallet[9:11])) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45)) +
    # labs(x = "", y = "", title = "B) Fishing ports approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot() +
    geom_sf(data = subset(grid_eez_fp_sf, index = "NA"), color = "white") +
    geom_sf(data = subset(grid_eez_fp_sf, index != "NA"), aes(color = group),size = 0.1, alpha = 0.3) +
    geom_sf(data = land_sf, aes(), fill = "grey50") +
    facet_wrap(~ group) +
    # geom_point(data = rename(ports_df, group = state_postal), aes(x = lon, y = lat), color = "#E6A0C4") +
    theme(legend.position = "top") +
    # scale_color_manual(values = state_pallet,
    scale_color_manual(values = c(state_pallet[3:7],state_pallet[9:11]),
                       # labels = unique(grid_eez_fp_sf$group)) +
                       labels = grid_eez_fp_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

name <- my_path("G", "AcrossBoundaries/Results/Figures", name = "fp_reg_units.png")

ggsave(name,
       p_fp,
       width = 9,
       height = 9)
